A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-09 survey.tex
Explaining clustering lables (openings)

% The paper headers
\markboth{TCIAIG ~Vol.~X, No.~Y, Month~Year}%
{??? \MakeLowercase{\textit{et al.}}: A Survey of RTS Game AI Research and Competition in StarCraft}

@@

\begin{abstract}
This paper presents an overview of the existing work on AI for real-time strategy (RTS) games. Specifically, we focus on the work around the game {\em StarCraft}, which has emerged in the past few years as the unified test-bed for this research. We describe the specific AI challenges posed by RTS games, and overview the solutions that have been explored to address them. Additionally, we also present a summary of the results of the recent StarCraft AI competitions, describing the architectures used by the participants. Finally, we conclude with a discussion emphasizing which problems in the context of RTS game AI have been solved, and which remain open.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
\IEEEPARstart{T}{he} field of real-time strategy (RTS) game AI has advanced significantly since Michael Buto's call for research in this area \cite{Buro03rts}. Specially, competitions like the ``AIIDE StarCraft AI Competition'' have motivated the exploration of many AI approaches in the context of RTS AI. We will list and classify these approaches, explain their 
strengths and weaknesses and conclude on what is left to achieve human-level 
RTS AI.

Complex dynamic environments, where it is not possible to have perfect nor complete information about the current state or about the dynamics of the environment, pose significant challenges for artificial intelligence. Road  traffic, finance, or
weather  forecasts are  examples of such large, complex, real-life dynamic
environments. RTS games can be seen as a simplification of a real-life
environment, with simpler dynamics in a  finite and smaller
world, although still complex enough  to bring very difficult and interesting
problems  to  the scientific  community like  decision making  under
uncertainty  or  real-time adversarial planning. Finding efficient  techniques for
tackling  these problems  on  RTS  games can  thus  benefit broader  AI
domains, but  also have concrete  and direct applications in  the ever growing
industry of  video games.%, generating worldwide two  and a  half times more revenue than the Film industry in 2011.
% This statement is not fair, the reference I found it only compare all the game industry revenue (including hardware sales) vs theatrical film industry revenue (i.e. not DVD sales, ...)

This paper, aiming to provide a one-stop guide on what is the state of
the art in RTS AI, and with a particular emphasis on the work done in StarCraft. 
It is organized as follows: Section~\ref{sec:rts}
introduces real-time strategy games, in particular the game StarCraft,
and  the main  AI challenges  in RTS  games.  Section~\ref{sec:review}
reviews  the existing  work on tackling these challenges in RTS games.
Section~\ref{sec:bot} analyzes several current state of the art RTS game playing agents (called {\em bots}), selected form the participants to annual
StarCraft    AI    competitions.
Section~\ref{sec:competition}  presents results of the recent annual competitions
held at the AIIDE  and CIG conferences and a StarCraft bot game ladder\footnote{An extended tournament, which can potentially go on indefinitely.}.  Section~\ref{sec:questions}
compiles open questions in RTS game AI. Finally, the paper concludes on discussions and perspectives.

@@

Similarly, current bots do not exhibit adaptation to opponent strategies. Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} show evidence of this. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). Specifically, we analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). As the figures show, the top three ranked bots in the competition (Skynet, Aiur and UalbertaBot) do not change their strategy at all depending on their opponent. For example, the Skynet bot (both in 2011 and 2012), always uses the same opening (``two gates''), except when playing a Terran opponent, when it uses ``nony''. This reflects the trend that the performance of bots is still more dependent on carefully hand-crafted and non-adaptive behaviors, than on on-line decision making procedures. This is so, since most of the problems that need to be solved in order to implement such procedures are still open.

@@

%\begin{IEEEbiography}[{\includegraphics[width=2cm, keepaspectratio]{figures/santi.jpg}}]{Santiago Onta\~{n}\'{o}n}

\begin{IEEEbiographynophoto}{Santiago Onta\~{n}\'{o}n}
is an assistant professor in the Computer Science Department at Drexel University. His main research interests are game AI, case-based reasoning and machine learning, fields in which he has published more than 90 peer-reviewed papers. He obtained his PhD form the Autonomous University of Barcelona (UAB), Spain. Before joining Drexel University, he held postdoctoral research positions at the Artificial Intelligence Research Institute (IIIA) in Barcelona, Spain, at the Georgia Institute of Technology (GeorgiaTech) in Atlanta, USA, and at the University of Barcelona, Spain.  
\end{IEEEbiographynophoto}

>>>>

% The paper headers
\markboth{TCIAIG ~Vol.~X, No.~Y, Month~Year}%
{Shell \MakeLowercase{\textit{et al.}}: TODO: Title here}

@@

\begin{abstract}
This paper presents an overview of the existing work on creating AI for real-time strategy (RTS) games. Specifically, we focus on the work performed around the game {\em StarCraft}, which has emerged in the past few years as the unified test-bed for this line of research. We describe the specific AI challenges posed by RTS games, and overview the solutions that have been explored to address them. Additionally, we also present a summary of the results of the recent StarCraft AI competitions, describing some of the architectures used by the participants. Finally, we conclude with a discussion emphasizing which problems in the context of RTS game AI have been solved, and which remain open.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
\IEEEPARstart{S}{ince} Michael Buro's call for research in real-time strategy (RTS) games \cite{Buro03rts}, many researchers have answered the call. Specially, AI competitions like the ``AIIDE StarCraft AI Competition'' have caused many AI techniques to be
applied to RTS AI. We will list and classify these approaches, explain their 
power and their downsides and conclude on what is left to achieve human-level 
RTS AI.

Many  challenging AI  problems can  be found  in  dynamic environments
where  one does  not have  a  perfect, complete  information of  their
current  state, as well  as their  evolution.  Road  traffic, finance,
weather  forecasts are  examples of  huge, complex,  real-life dynamic
environments. RTS games can be seen as a simplification of a real-life
environment, where there  exist simpler rules in a  finite and smaller
world, but are complex enough  to bring very difficult and interesting
problems  to  the scientific  community,  like  decision making  under
uncertainty  or  real-time adversarial planning.   Finding  efficient  techniques
to address these problems  on  RTS  games can  thus  profit broader  AI
domains, but  have also concrete  and direct applications in  the huge
industry of  video games.
%,  generating worldwide two  and a  half times more revenue than the Film industry in 2011.
% This statement is not fair, the reference I found it only compare all the game industry revenue (including hardware sales) vs theatrical film industry revenue (i.e. not DVD sales, ...)

This paper, aiming to provide a one-stop guide on what is the state of
the art in RTS AI, and with a particular emphasis on the work done in StarCraft. 
It is organized as follows: Section~\ref{sec:rts}
introduces real-time strategy games, in particular the game StarCraft,
and  the main  AI challenges  in RTS  games.  Section~\ref{sec:review}
reviews  the existing  work on  RTS games  tackling  these challenges.
Section~\ref{sec:bot} shows  the current state  of the art of bots, i.e. a computer-controlled players, 
for  StarCraft, presenting  seven bots participating  to annual
StarCraft    AI    competitions     for    a    couple    of    years.
Section~\ref{sec:competition}  presents these two  annual competitions
held  at the AIIDE  and CIG  conferences and a StarCraft bot game ladder\footnote{An extended tournament, which can potentially go on indefinitely.}.  Section~\ref{sec:questions}
compiles open questions  in RTS game AI which must  be tackled to make
bots reaching a high human-level play.
% skip "empirical evidence..." for the moment
Finally, the paper concludes on discussions and perspectives.

@@

Similarly, current bots do not exhibit adaptation to opponent strategies. Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} show evidence of this. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). Specifically, we analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). Now we give a description of the different openings that we considered to build the clustering. Most of the following names are specific building or units of StarCraft, for a better comprehension we refer the reader to \footnote{\url{http://wiki.teamliquid.net/starcraft/Category:Strategies}}:
\begin{itemize}
  \item Protoss openings:
  \begin{itemize}
    \item \emph{two\_gates}: Build two \emph{Gateways} and keep training \emph{Zealots}.
    \item \emph{fast\_dt}: Produce \emph{Dark Templars} as soon as possible.
    \item \emph{templar}: Train \emph{High Templars} as fast as possible.
    \item \emph{speedzeal}: Train \emph{Zealots}  and research attack and speed upgrades as soon as possible.
    \item \emph{corsair}: Produce \emph{Corsairs} as soon as possible and then make transition to train \emph{Dark Templars}, \emph{Reavers} or \emph{Dragoons}.
    \item \emph{nony}: Build three \emph{Gateways} and massive training of \emph{Dragoons}.
    \item \emph{reaver\_drop}: Train \emph{Reavers} as soon as possible to be able to do drops.
  \end{itemize}
  \item Terran openings:
  \begin{itemize}
    \item \emph{bio}: Produce a large army of \emph{Marines} and \emph{Medics}.
    \item \emph{rax\_fe}: Take the closest ``natural expansion'' as soon as possible. This provides a big economic boost in the early to mid game.
    \item \emph{two\_facto}: Build two \emph{Factories} and keep producing \emph{Tanks}.
    \item \emph{vultures}: Produce mainly \emph{Vultures} and research mines.
    \item \emph{air}: Produce \emph{Wraiths} as soon as possible for an air attack.
    \item \emph{drop}: Train \emph{Dropships} as soon as possible to be able to do drops. 
  \end{itemize}
  \item Zerg openings:
  \begin{itemize}
    \item \emph{speedlings}: Train \emph{Zerlings} and research speed upgrade as soon as possible.
    \item \emph{fast\_mutas}: Produce mainly \emph{Mutalisks}.
    \item \emph{mutas}: Expand two times for a stronger economy before massive training of \emph{Mutalisks}.
    \item \emph{lurkers}: Train \emph{Lurkers} as soon as possible to take benefit of its cloak ability.
    \item \emph{hydras}: Massive production of \emph{Hydralisks}.
  \end{itemize}
\end{itemize}
As the figures show, the top three ranked bots in the competition (Skynet, Aiur and UalbertaBot) do not change their strategy at all depending on their opponent. For example, the Skynet bot (both in 2011 and 2012), always uses the same opening (\emph{two\_gates}), except when playing a Terran opponent, when it uses \emph{nony}. This reflects the trend that the performance of bots is still more dependent on carefully hand-crafted and non-adaptive behaviors, than on on-line decision making procedures. This is so, since most of the problems that need to be solved in order to implement such procedures are still open.

@@

%\begin{IEEEbiography}[{\includegraphics[width=2cm, keepaspectratio]{figures/santi.jpg}}]{Santiago Onta\~{n}\'{o}n}
\begin{IEEEbiographynophoto}{Santiago Onta\~{n}\'{o}n}
is an assistant professor in the Computer Science Department at Drexel University. His main research interests are game AI, case-based reasoning and machine learning, fields in which he has published more than 90 peer-reviewed papers. He obtained his PhD form the Autonomous University of Barcelona (UAB), Spain. Before joining Drexel University, he held postdoctoral research positions at the Artificial Intelligence Research Institute (IIIA) in Barcelona, Spain, at the Georgia Institute of Technology (GeorgiaTech) in Atlanta, USA, and at the University of Barcelona, Spain.  
\end{IEEEbiographynophoto}
